### 2024-06-03 main.tex
final

\newcommand{\N}{\mathbf{N}}
\newcommand{\twoj}{\nu}
\newcommand{\fra}[1]{\textcolor[rgb]{0,0,1}{#1}}
\allowdisplaybreaks

\begin{document} 
 \linenumbers
\begingroup

@@

\paragraph{The model.}
The model we consider is the {\em  multispecies stirring process with open boundaries}, 
i.e. the multispecies analogue of the exclusion process as introduced in \cite{zhou2021orthogonal}, 
with additional boundary reservoirs as proposed in \cite{vanicat2017exact}. 
The hydrodynamic limit and fluctuations of the multispecies stirring process on $\mathbb{Z}^d$
has been recently studied in \cite{casini2023density}.
In our set-up, particles can be of $N-1$ types (or species) and they can move
on a generic connected and undirected graph $G=(V,\mathcal{E})$ with vertex set $V$ and
edge set $\mathcal{E}$. The maximal number of particles allowed at each vertex (also called \textit{maximal occupancy}) is denoted by $\nu\in \mathbb{N}$. 
Two particles do interact when they sit on two sites connected by a graph edge and this interaction consists in swapping them. 
Furthermore, reservoirs are attached at each site of the graph, creating and removing particles. For a precise definition of the rates of the process 
we refer to Section~\ref{sec1}. As we will see, the choice of the boundary reservoirs is motivated by the fact that they allow to determine an absorbing dual process
\fra{that does not permit  color-changing until absorption at the boundary}, 
regardless of the graph $G$ on which the process is defined and of the maximal occupancy $\nu$ at each site.

@@

\paragraph{Organization of the paper and main results.}
We start in Section~\ref{sec1} by \fra{defining} the open multispecies stirring
process on a general graph and providing its Lie algebraic description in terms of the generators of the $gl(N)$ Lie algebra. In Section~\ref{sectionDuality} we prove an absorption duality (Theorem~\ref{thm-duality})
and show that the problem of characterizing the non-equilibrium
steady state can be reduced to the computation of the absorption probabilities similar to the single species case on the line  \cite{KMP} and for a general graph \cite{ frassek2020duality}.  
In Section~\ref{sectionIntegrabiliy}  we consider  the case of the one-dimensional chain with at most one 
particle per site ($\nu=1$), which is driven out-of-equilbrium by two reservoirs on the left and right boundaries. In this case we obtain a closed-form expression for the stationary correlations (Theorem~\ref{thm-correlations}). Here the computation
relies on the extension of the  Matrix Product Ansatz \cite{derrida1993exact} to multiple species \cite{vanicat2017exact} and the $gl(N)$ invariance of the process in the bulk. Finally in Section~\ref{sec5} we discuss a few extensions 
and generalizations of the model. 
The first generalization is the formula for the correlations for the boundary driven chain in the case when more than one particle is allowed at each site ($\nu > 1$). 
In this situation integrability is lost and we cannot derive explicit formulas for the absorption probabilities. However, in Appendix \ref{appendix-1pts-general}, we show how to compute the one-point correlation by using the properties of a random walk. The second generalization is the definition of a thermalized version of the boundary-driven multispecies stirring process and the formulation of the the absorbing duality with the same duality function of the multispecies stirring process (the details of the proof are reported in Appendix \ref{appendix-dualityThermalized}). \fra{Finally, inspired by \cite{casini2022uphill,schutzReaction},} the third generalization is the introduction of an ``ad-hoc'' interaction mechanism of reaction type, in order to have closed evolution equation for average particle densities, \fra{(we report the details of the computations for the closure of the average density evolution equation in Appendix \ref{appendix-RD})}. 
Again, an absorbing duality is proved with the same duality function of the multispecies stirring process.

@@

\section{The stirring process with open boundaries}
\label{sec1}
\subsection{Informal description of the process}
The process studied in this paper is the {\em multispecies stirring process}
with open boundaries. 
Each site
of a connected graph can host a maximal number of particles called $\nu\in \mathbb{N}$ .
The particles have a {\em type} (sometimes called {\em species} or {\em colour})
which can \fra{take}  values $\{1,2,\ldots,N-1\}$. {More precisely, at each site there are $\nu$ available places (called \textit{holes} or \textit{vacancies}) that can be either occupied by particles or not. As a consequence, each site can be totally empty (when no particles are present and thus $\nu$ holes are available), totally occupied (when $\nu$ particles, of any type, are present and thus no holes are available) or partially occupied (when some of the holes are occupied and some are available). All these configurations obey the \textit{exclusion constraint} that tells that, at each site and at any time, the sum of the number of particles of any type plus the number of holes is $\nu$.}

Then, the dynamics has two parts: 
\begin{itemize}
\item
on each edge of the graph, any two types of particles are swapped at rate $1$;
moreover a particle of any type and a hole are also swapped at rate $1$; 
\item
on each vertex $x$ of the graph, a particle of type $a \in \{1,\ldots,N-1\}$ is created  at rate $\alpha_a^x >0$;
additionally a \fra{particle} of any type is replaced with a
particle of type $a$ at rate $\alpha_a^x >0$ and it is removed at rate $\alpha_N^x >0$.
\end{itemize}
The swap dynamics taking place on the edges is of Kawasaki-type 
with $N$ conservation laws, \fra{where $N$ is the number of species plus the holes. This means that the sum of the particle number of each species and the number of holes is conserved by the dynamics.}
%(the total number of particles of each type). 
The site-dynamics is instead of Glauber-type. 
In the long-time limit, a so-called non-equilibrium
steady state sets in.
In the case $N=2$, we retrieve the  boundary-driven version of the symmetric exclusion process \cite{schutzSandow,carinci2013duality}.

@@

Namely, on each site $x\in V$  a particle {or a hole} indicated by $B$ is replaced with a particle {or a hole} indicated $A$ at rate $\Gamma_{x}\alpha_{A}^{x}n_{B}^{x}$. {The transitions where \fra{a} particle is replaced by a hole/a hole is replaced by a particle is interpreted as a removal/injection of particles.}

@@

and  $|\alpha^x|=\sum_{A=1}^{N}\alpha_{A}^x$. {As a consequence, particles are injected and removed in the graph with different rates. When, for any $A\in \{1,\ldots,N\}$, there are at least two reservoirs at two different sites $x,y$ such that $\rho_{A}^{x}\neq  \rho_{A}^{y}$, a  non-zero current sets in, breaking the reversibility. Informally, one can say that the reservoirs try to impose their different densities at the boundaries of \fra{the} graph, putting the system out-of-equilibrium.} 

@@

For  $\nu=1$ we denote the ${gl}(N)$  generators   as  $(e_{AB})_{A,B\in\{1,\ldots,N\}}$ obeying $(e_{AB})_{CD}=\delta_{AC}\delta_{BD}$ \footnote{\fra{In this set-up we have chosen the fundamental (defining) representation of the $gl(N)$ Lie algebra, therefore the generators $(E_{AB})_{A,B\in\{1,\ldots,N\}}$ do coincide with the elementary matrices $(e_{AB})_{A,B\in\{1,\ldots,N\}}$. As a consequence, we denote these generators by lowercase letters}}.

@@

Under these assumption the Hamiltonian~\ref{hamiltonian} is integrable. This has been \fra{shown} in \cite{vanicat2017exact} within  the Quantum Inverse Scattering Method.  We will consider this process more extensively in Section~\ref{sectionIntegrabiliy}.

@@

$\mathcal{L}_{x,y}$ \fra{are} defined in \eqref{edgeGenerator} and, for any function $f:\widetilde{\Omega}\to \mathbb{R}$ 

@@

	\langle q_{1}^{u(x)},\ldots,q^{u(x)}_{N-1}|\xi_{1}^{u(x)},\ldots,\xi_{N-1}^{u(x)}\rangle=\prod_{a=1}^{N-1}\delta_{\xi_{a}^{u(x)},\fra{q_{a}^{u(x)}}}\,.

@@

Since \eqref{dualityMatrix} is product over sites, proving \eqref{DualityRelation} is equivalent to showing that 
\begin{equation}\label{edgeDualRealtion}
    \mathcal{H}_{x,y}^{T}D=D\mathcal{H}_{x,y}\qquad \forall (x,y)\in \mathcal{E}
\end{equation}
and 
\begin{equation}\label{siteDualRelation}
    H_{x}^{T}D=D\widetilde{H}_{x}\qquad \forall x\in V.
\end{equation}
We perform the proof of duality in three steps: first we will show that matrix \eqref{dualityMatrix} has elements \eqref{dualityElements}; second we will prove the bulk duality \eqref{edgeDualRealtion}; finally we will show the boundary duality \eqref{siteDualRelation}. 
As a preliminary result we note that 
\begin{equation}\label{transpositionPropertyR}
(E_{BA}^{x})^{T}=R_{x}E_{AB}^{x}R_{x}^{-1}\qquad \forall x\in V\co
\end{equation} 
that follows immediately from the definition of $E_{AB}$. 
\paragraph{Elements of the duality matrix.}
Consider the matrix $D$ defined in \eqref{dualityMatrix}.
We aim to show that its matrix elements coincide with duality function \eqref{dualityElements}, i.e.
\begin{equation}\label{proofDualityElements}
\langle \bm{n}|D|\bm{\xi}\rangle=D(\bm{n},\bm{\xi})\qquad   \forall \bm{n}\in \Omega,\quad \bm{\xi}\in \widetilde{\Omega}\,.
\end{equation}
Fix an arbitrary site $x\in V$, then we have that 
\begin{align}
	 &\langle n^{x}_{1},\ldots,n_{N}^{x}|\left(d_{x}\otimes \dd\right)\left(|\xi_{1}^{x},\ldots,\xi_{N}^{x}\rangle\otimes|\xi_{1}^{u(x)},\ldots,\xi_{N-1}^{u(x)}\rangle\right)\nonumber\\=&\langle n_{1}^{x},\ldots,n_{N}^{x}| (\exp{(E_{N1}^{x}+\ldots+E_{N\,N-1}^{x}}))^{T}R_{x}\otimes\sum_{q_{1}^{u(x)},\ldots,q_{N-1}^{u(x)}=0}^{\infty}\prod_{a=1}^{N-1}\left(\rho_{a}^{x}\right)^{q_{a}^{u(x)}}\langle q_{1}^{u(x)},\ldots,q_{N}^{u(x)}|\nonumber
	 \\&|\xi_{1}^{x},\ldots,\xi_{N}^{x}\rangle \otimes |\xi_{1}^{u(x)},\ldots,\xi_{N-1}^{u(x)}\rangle
\end{align} 
where we used \eqref{transpositionPropertyR}. \\
On one hand, on the extra-site $u(x)$ we have 
\begin{align}
\sum_{q_{1}^{u(x)},\ldots,q_{N-1}^{u(x)}=0}^{\infty}\left(\prod_{a=1}^{N-1}\left(\rho_{a}^{x}\right)^{q_{a}^{u(x)}}\right)\langle q_{1}^{u(x)},\ldots,q_{N-1}^{u(x)}|\xi_{1}^{u(x)},\ldots,\xi_{N-1}^{u(x)}\rangle=\prod_{a=1}^{N-1}\left(\rho_{a}^{x}\right)^{\xi_{a}^{u(x)}}
\end{align}
{ 
where we used the orthogonality relation \eqref{ortho-extraSite}}. 
On the other hand, on the site $x$, we have 
\begin{align}
&\langle n_{1}^{x},\ldots,n_{N}^{x}|(\exp{(E_{N1}^{x}+\ldots+E_{N\,N-1}^{x})})^{T}R_{x}|\xi_{1}^{x},\ldots,\xi_{N}^{x}\rangle\nonumber\\&= \langle  n_{1}^{x},\ldots,n_{N}^{x}|\left(\sum_{k_{1}=0}^{\infty}\frac{\left\{\left(E_{N1}^{x}\right)^{T}\right\}^{k_{1}}}{k_{1}!}\ldots\sum_{k_{N-1}=0}^{\infty}\frac{\left\{\left(E_{N\,N-1}^{x}\right)^{T}\right\}^{k_{N-1}}}{k_{N-1}!}\fra{\sum_{r\in\Omega_{x}}\frac{r_{1}^{x}!\cdots r_{N}^{x}!}{\nu!}|r_{1}^{x},\ldots,r_{N}^{x}\rangle\langle r_{1}^{x},\ldots,r_{N}^{x}|}\right)\nonumber
\\&
\hspace{2.cm}\qquad\times |\xi_{1}^{x},\ldots,\xi_{N}^{x}\rangle\nonumber
\\&=
\sum_{k_{1}=0}^{n_{1}^{x}}\ldots\sum_{k_{N-1}=0}^{n_{N-1}^{x}}\langle n_{1}^{x}-k_{1},\ldots,n_{N-1}^{x}-k_{N-1},\ldots,n_{N}^{x}+k_{1}+\ldots+k_{N-1}|\frac{n_{1}^{x}!\cdots n_{N-1}^{x}!}{(n_{1}^{x}-k_{1})!\cdots(n_{N-1}^{x}-k_{N-1})!}\nonumber
\\& 
\hspace{2.cm}\qquad\times \frac{1}{k_{1}!\cdots k_{N-1}!}\frac{\xi_{1}^{x}!\cdots \xi_{N}^{x}!}{\nu!}|\xi_{1}^{x},\ldots,\xi_{N}^{x}\rangle\nonumber
\\&=
\frac{(\nu-\sum_{a=1}^{N-1}\xi_{a}^{x})!}{\nu!}\prod_{a=1}^{N-1}\frac{n_{a}^{x}!}{(n_{a}^{x}-\xi_{a}^{x})!}
\end{align}
where we used the definitions of the action of $E_{AB}$  and the orthogonality relations \eqref{ortho}, together with the fact that $\xi_{N}^{x}=\nu-\sum_{a=1}^{N-1}\xi_{a}^{x}$. 
Finally, by taking the product over $x\in V$ \eqref{proofDualityElements} is proved.
\begin{flushright}
    $\square$
\end{flushright}

@@

 \paragraph{Proof of boundary duality \eqref{siteDualRelation}.} To prove \eqref{siteDualRelation} we transform via the Hadamard formula \footnote{\fra{See \cite{khuan} for an application of the Hadamard formula in the multispecies asymmetric context.}} the transposed of the site Hamiltonian \eqref{siteDualHamiltonian} and then we  introduce a creation operator acting on an extra-site $u(x)$.

@@

 	\fra{\exp{\left(%-\sum_{a=1}^{N-1}E_{aN}

@@

\fra{It follows that}

@@

	\fra{\exp{(-E)}\left(\sum_{a,b=1\,b\neq a}^{N-1}\alpha_{a}\left(E_{ba}-E_{bb}\right)\right)\exp{(E)}=\sum_{a,b=1\,b\neq a}^{N-1}\alpha_{a}\left(E_{ba}-E_{bb}\right)}\co

@@

	\fra{\exp{(-E)}\left(\sum_{a=1}^{N-1}\alpha_{a}\left(E_{Na}-E_{NN}\right)\right)\exp{(E)}=\sum_{a=1}^{N-1}\left(E_{Na}-E_{aa}-\sum_{b=1\,b\neq a}^{N-1}E_{ba}\right)}
\fra{and}

@@

	\fra{\exp{(-E)}\left(\alpha_{N}\sum_{b=1}^{N-1}\alpha_{a}\left(E_{bN}-E_{bb}\right)\right)\exp{(E)}=-\alpha_{N}\sum_{b=1}^{N-1}E_{bb}}

@@

By using \fra{\eqref{uno},\eqref{due} and \eqref{tre})} and the fact that 
\fra{$\sum_{A,B=1}^{N}E_{AB}$} is central for the algebra we have

@@

\fra{\sum_{a=1}^{N-1}\alpha_{a}\left(E_{Na}^{x}-E_{aa}^{x}\right)-\sum_{a,b=1\, a\neq b}^{N-1}\alpha_{a}^{x}E_{bb}^{x}-\alpha_{N}\sum_{a=1}^{N-1}E_{aa}^{x}\nonumber}

@@

\fra{	|\alpha^{x}|\sum_{a=1}^{N-1}\left(\rho_{a}^{x}E_{Na}^{x}-E_{aa}^{x}\right)}\dt

@@

we obtain \fra{

@@

{With a slight abuse of notation and in the spirit of Remark \ref{reamrk-extraSite}, we here denote the occupation vector at extra-site $u(x)$ by $\langle \xi_{1}^{u(x)},\ldots,\xi_{N-1}^{u(x)}|$\fra{.}

@@

\fra{Recalling} the action of the bosonic creation operator acting at site $u(x)$ and on the species $a\in \{1,\ldots,N-1\}$ we have that 

@@

Using the above equation \fra{and considering the addends multiplied by the Lie generator $E_{Na}^{x}$ (for all $a\in \{1,\ldots,N-1\}$) on the right hand side term of \eqref{siteHadamardII}} we have that 

@@

In the long-time limit the dual process voids the chain, i.e. all particles of types $\{1,\ldots,N-1\}$ are eventually absorbed at the extra-sites $0$ and $L+1$ and replaced by types $N$. We also notice that, \fra{up to the the bosonic} creation operators $(\mathbf{a}^{+})_{a}$, the dual boundary Hamiltonians \eqref{boundary-H-dual-chain} are triangular. This simplification will be crucial in Section~\ref{sectionIntegrabiliy}.\\

@@

\section{Correlations and steady state  for the integrable chain}\label{sectionIntegrabiliy}
In this section we combine the matrix product ansatz (MPA) \cite{vanicat2017exact} and  the $gl(N)$ invariance of the process in the bulk, cf.~\eqref{symmetryH}, to exactly compute the correlations in the non-equilibrium steady state
 for the integrable version of the multispecies stirring process \fra{defined on a line segment}, see Section~\ref{integrableChain-duality}. 
This allows further to determine the absorption probabilities and the probability mass-function of the non-equilibrium steady state in closed form.
We start by recalling the formulation of the MPA  and then we explain our strategy.

>>>>

\newcommand{\N}{\mathbf{N}}
\newcommand{\twoj}{\nu}
\newcommand{\fra}[1]{\textcolor[rgb]{1,0,0}{#1}}
\newcommand{\fraT}[1]{\textcolor[rgb]{0,0,0}{#1}}
\allowdisplaybreaks

\begin{document} 
 %\linenumbers
\begingroup

@@

\paragraph{The model.}
The model we consider is the {\em  multispecies stirring process with open boundaries}, 
i.e. the multispecies analogue of the exclusion process as introduced in \cite{zhou2021orthogonal}, 
with additional boundary reservoirs as proposed in \cite{vanicat2017exact}. 
The hydrodynamic limit and fluctuations of the multispecies stirring process on $\mathbb{Z}^d$
has been recently studied in \cite{casini2023density}.
In our set-up, particles can be of $N-1$ types (or species) and they can move
on a generic connected and undirected graph $G=(V,\mathcal{E})$ with vertex set $V$ and
edge set $\mathcal{E}$. The maximal number of particles allowed at each vertex (also called \textit{maximal occupancy}) is denoted by $\nu\in \mathbb{N}$. 
Two particles do interact when they sit on two sites connected by a graph edge and this interaction consists in swapping them. 
Furthermore, reservoirs are attached at each site of the graph, creating and removing particles. For a precise definition of the rates of the process 
we refer to Section~\ref{sec1}. As we will see, the choice of the boundary reservoirs is motivated by the fact that they allow to determine an absorbing dual process
\fraT{that does not permit  color-changing until absorption at the boundary}, 
regardless of the graph $G$ on which the process is defined and of the maximal occupancy $\nu$ at each site.

@@

\paragraph{Organization of the paper and main results.}
We start in Section~\ref{sec1} by \fraT{defining} the open multispecies stirring
process on a general graph and providing its Lie algebraic description in terms of the generators of the $gl(N)$ Lie algebra. In Section~\ref{sectionDuality} we prove an absorption duality (Theorem~\ref{thm-duality})
and show that the problem of characterizing the non-equilibrium
steady state can be reduced to the computation of the absorption probabilities similar to the single species case on the line  \cite{KMP} and for a general graph \cite{ frassek2020duality}.  
In Section~\ref{sectionIntegrabiliy}  we consider  the case of the one-dimensional chain with at most one 
particle per site ($\nu=1$), which is driven out-of-equilbrium by two reservoirs on the left and right boundaries. In this case we obtain a closed-form expression for the stationary correlations (Theorem~\ref{thm-correlations}). Here the computation
relies on the extension of the  Matrix Product Ansatz \cite{derrida1993exact} to multiple species \cite{vanicat2017exact} and the $gl(N)$ invariance of the process in the bulk. Finally in Section~\ref{sec5} we discuss a few extensions 
and generalizations of the model. 
The first generalization is the formula for the correlations for the boundary driven chain in the case when more than one particle is allowed at each site ($\nu > 1$). 
In this situation integrability is lost and we cannot derive explicit formulas for the absorption probabilities. However, in Appendix \ref{appendix-1pts-general}, we show how to compute the one-point correlation by using the properties of a random walk. The second generalization is the definition of a thermalized version of the boundary-driven multispecies stirring process and the formulation of the the absorbing duality with the same duality function of the multispecies stirring process (the details of the proof are reported in Appendix \ref{appendix-dualityThermalized}). \fraT{Finally, inspired by \cite{casini2022uphill,schutzReaction},} the third generalization is the introduction of an ``ad-hoc'' interaction mechanism of reaction type, in order to have closed evolution equation for average particle densities, \fraT{(we report the details of the computations for the closure of the average density evolution equation in Appendix \ref{appendix-RD})}. 
Again, an absorbing duality is proved with the same duality function of the multispecies stirring process.

@@

\section{The stirring process with open boundaries}
\label{sec1}
\subsection{Informal description of the process}
The process studied in this paper is the {\em multispecies stirring process}
with open boundaries. 
Each site
of a connected graph can host a maximal number of particles called $\nu\in \mathbb{N}$ .
The particles have a {\em type} (sometimes called {\em species} or {\em colour})
which can \fraT{take}  values $\{1,2,\ldots,N-1\}$. {More precisely, at each site there are $\nu$ available places (called \textit{holes} or \textit{vacancies}) that can be either occupied by particles or not. As a consequence, each site can be totally empty (when no particles are present and thus $\nu$ holes are available), totally occupied (when $\nu$ particles, of any type, are present and thus no holes are available) or partially occupied (when some of the holes are occupied and some are available). All these configurations obey the \textit{exclusion constraint} that tells that, at each site and at any time, the sum of the number of particles of any type plus the number of holes is $\nu$.}

Then, the dynamics has two parts: 
\begin{itemize}
\item
on each edge of the graph, any two types of particles are swapped at rate $1$;
moreover a particle of any type and a hole are also swapped at rate $1$; 
\item
on each vertex $x$ of the graph, a particle of type $a \in \{1,\ldots,N-1\}$ is created  at rate $\alpha_a^x >0$;
additionally a \fraT{particle} of any type is replaced with a
particle of type $a$ at rate $\alpha_a^x >0$ and it is removed at rate $\alpha_N^x >0$.
\end{itemize}
The swap dynamics taking place on the edges is of Kawasaki-type 
with $N$ conservation laws, \fraT{where $N$ is the number of species plus the holes. This means that the sum of the particle number of each species and the number of holes is conserved by the dynamics.}
%(the total number of particles of each type). 
The site-dynamics is instead of Glauber-type. 
In the long-time limit, a so-called non-equilibrium
steady state sets in.
In the case $N=2$, we retrieve the  boundary-driven version of the symmetric exclusion process \cite{schutzSandow,carinci2013duality}.

@@

Namely, on each site $x\in V$  a particle {or a hole} indicated by $B$ is replaced with a particle {or a hole} indicated $A$ at rate $\Gamma_{x}\alpha_{A}^{x}n_{B}^{x}$. {The transitions where \fraT{a} particle is replaced by a hole/a hole is replaced by a particle is interpreted as a removal/injection of particles.}

@@

and  $|\alpha^x|=\sum_{A=1}^{N}\alpha_{A}^x$. {As a consequence, particles are injected and removed in the graph with different rates. When, for any $A\in \{1,\ldots,N\}$, there are at least two reservoirs at two different sites $x,y$ such that $\rho_{A}^{x}\neq  \rho_{A}^{y}$, a  non-zero current sets in, breaking the reversibility. Informally, one can say that the reservoirs try to impose their different densities at the boundaries of \fraT{the} graph, putting the system out-of-equilibrium.} 

@@

For  $\nu=1$ we denote the ${gl}(N)$  generators   as  $(e_{AB})_{A,B\in\{1,\ldots,N\}}$ obeying $(e_{AB})_{CD}=\delta_{AC}\delta_{BD}$ \footnote{\fraT{In this set-up we have chosen the fundamental (defining) representation of the $gl(N)$ Lie algebra, therefore the generators $(E_{AB})_{A,B\in\{1,\ldots,N\}}$ do coincide with the elementary matrices $(e_{AB})_{A,B\in\{1,\ldots,N\}}$. As a consequence, we denote these generators by lowercase letters}}.

@@

Under these assumption the Hamiltonian~\ref{hamiltonian} is integrable. This has been \fraT{shown} in \cite{vanicat2017exact} within  the Quantum Inverse Scattering Method.  We will consider this process more extensively in Section~\ref{sectionIntegrabiliy}.

@@

$\mathcal{L}_{x,y}$ \fraT{are} defined in \eqref{edgeGenerator} and, for any function $f:\widetilde{\Omega}\to \mathbb{R}$ 

@@

	\langle q_{1}^{u(x)},\ldots,q^{u(x)}_{N-1}|\xi_{1}^{u(x)},\ldots,\xi_{N-1}^{u(x)}\rangle=\prod_{a=1}^{N-1}\delta_{\xi_{a}^{u(x)},\fraT{q_{a}^{u(x)}}}\,.

@@

Since \eqref{dualityMatrix} is product over sites, proving \eqref{DualityRelation} is equivalent to showing that 
\begin{equation}\label{edgeDualRealtion}
    \mathcal{H}_{x,y}^{T}D=D\mathcal{H}_{x,y}\qquad \forall (x,y)\in \mathcal{E}
\end{equation}
and 
\begin{equation}\label{siteDualRelation}
    H_{x}^{T}D=D\widetilde{H}_{x}\qquad \forall x\in V.
\end{equation}
We perform the proof of duality in three steps: first we will show that matrix \eqref{dualityMatrix} has elements \eqref{dualityElements}; second we will prove the bulk duality \eqref{edgeDualRealtion}; finally we will show the boundary duality \eqref{siteDualRelation}. 
As a preliminary result we note that 
\begin{equation}\label{transpositionPropertyR}
(E_{BA}^{x})^{T}=R_{x}E_{AB}^{x}R_{x}^{-1}\qquad \forall x\in V\co
\end{equation} 
that follows immediately from the definition of $E_{AB}$. 
\paragraph{Elements of the duality matrix.}
Consider the matrix $D$ defined in \eqref{dualityMatrix}.
We aim to show that its matrix elements coincide with duality function \eqref{dualityElements}, i.e.
\begin{equation}\label{proofDualityElements}
\langle \bm{n}|D|\bm{\xi}\rangle=D(\bm{n},\bm{\xi})\qquad   \forall \bm{n}\in \Omega,\quad \bm{\xi}\in \widetilde{\Omega}\,.
\end{equation}
Fix an arbitrary site $x\in V$, then we have that 
\begin{align}
	 &\langle n^{x}_{1},\ldots,n_{N}^{x}|\left(d_{x}\otimes \dd\right)\left(|\xi_{1}^{x},\ldots,\xi_{N}^{x}\rangle\otimes|\xi_{1}^{u(x)},\ldots,\xi_{N-1}^{u(x)}\rangle\right)\nonumber\\=&\langle n_{1}^{x},\ldots,n_{N}^{x}| (\exp{(E_{N1}^{x}+\ldots+E_{N\,N-1}^{x}}))^{T}R_{x}\otimes\sum_{q_{1}^{u(x)},\ldots,q_{N-1}^{u(x)}=0}^{\infty}\prod_{a=1}^{N-1}\left(\rho_{a}^{x}\right)^{q_{a}^{u(x)}}\langle q_{1}^{u(x)},\ldots,q_{N}^{u(x)}|\nonumber
	 \\&|\xi_{1}^{x},\ldots,\xi_{N}^{x}\rangle \otimes |\xi_{1}^{u(x)},\ldots,\xi_{N-1}^{u(x)}\rangle
\end{align} 
where we used \eqref{transpositionPropertyR}. \\
On one hand, on the extra-site $u(x)$ we have 
\begin{align}
\sum_{q_{1}^{u(x)},\ldots,q_{N-1}^{u(x)}=0}^{\infty}\left(\prod_{a=1}^{N-1}\left(\rho_{a}^{x}\right)^{q_{a}^{u(x)}}\right)\langle q_{1}^{u(x)},\ldots,q_{N-1}^{u(x)}|\xi_{1}^{u(x)},\ldots,\xi_{N-1}^{u(x)}\rangle=\prod_{a=1}^{N-1}\left(\rho_{a}^{x}\right)^{\xi_{a}^{u(x)}}
\end{align}
{ 
where we used the orthogonality relation \eqref{ortho-extraSite}}. 
On the other hand, on the site $x$, we have 
\begin{align}
&\langle n_{1}^{x},\ldots,n_{N}^{x}|(\exp{(E_{N1}^{x}+\ldots+E_{N\,N-1}^{x})})^{T}R_{x}|\xi_{1}^{x},\ldots,\xi_{N}^{x}\rangle\nonumber\\&= \langle  n_{1}^{x},\ldots,n_{N}^{x}|\left(\sum_{k_{1}=0}^{\infty}\frac{\left\{\left(E_{N1}^{x}\right)^{T}\right\}^{k_{1}}}{k_{1}!}\ldots\sum_{k_{N-1}=0}^{\infty}\frac{\left\{\left(E_{N\,N-1}^{x}\right)^{T}\right\}^{k_{N-1}}}{k_{N-1}!}\fraT{\sum_{r\in\Omega_{x}}\frac{r_{1}^{x}!\cdots r_{N}^{x}!}{\nu!}|r_{1}^{x},\ldots,r_{N}^{x}\rangle\langle r_{1}^{x},\ldots,r_{N}^{x}|}\right)\nonumber
\\&
\hspace{2.cm}\qquad\times |\xi_{1}^{x},\ldots,\xi_{N}^{x}\rangle\nonumber
\\&=
\sum_{k_{1}=0}^{n_{1}^{x}}\ldots\sum_{k_{N-1}=0}^{n_{N-1}^{x}}\langle n_{1}^{x}-k_{1},\ldots,n_{N-1}^{x}-k_{N-1},\ldots,n_{N}^{x}+k_{1}+\ldots+k_{N-1}|\frac{n_{1}^{x}!\cdots n_{N-1}^{x}!}{(n_{1}^{x}-k_{1})!\cdots(n_{N-1}^{x}-k_{N-1})!}\nonumber
\\& 
\hspace{2.cm}\qquad\times \frac{1}{k_{1}!\cdots k_{N-1}!}\frac{\xi_{1}^{x}!\cdots \xi_{N}^{x}!}{\nu!}|\xi_{1}^{x},\ldots,\xi_{N}^{x}\rangle\nonumber
\\&=
\frac{(\nu-\sum_{a=1}^{N-1}\xi_{a}^{x})!}{\nu!}\prod_{a=1}^{N-1}\frac{n_{a}^{x}!}{(n_{a}^{x}-\xi_{a}^{x})!}
\end{align}
where we used the definitions of the action of $E_{AB}$  and the orthogonality relations \eqref{ortho}, together with the fact that $\xi_{N}^{x}=\nu-\sum_{a=1}^{N-1}\xi_{a}^{x}$. 
Finally, by taking the product over $x\in V$ \eqref{proofDualityElements} is proved.
\begin{flushright}
    $\square$
\end{flushright}

@@

 \paragraph{Proof of boundary duality \eqref{siteDualRelation}.} To prove \eqref{siteDualRelation} we transform via the Hadamard formula \footnote{\fraT{See \cite{khuan} for an application of the Hadamard formula in the multispecies asymmetric context.}} the transposed of the site Hamiltonian \eqref{siteDualHamiltonian} and then we  introduce a creation operator acting on an extra-site $u(x)$.

@@

 	\fraT{\exp{\left(%-\sum_{a=1}^{N-1}E_{aN}

@@

\fraT{It follows that}

@@

	\fraT{\exp{(-E)}\left(\sum_{a,b=1\,b\neq a}^{N-1}\alpha_{a}\left(E_{ba}-E_{bb}\right)\right)\exp{(E)}=\sum_{a,b=1\,b\neq a}^{N-1}\alpha_{a}\left(E_{ba}-E_{bb}\right)}\co

@@

	\fraT{\exp{(-E)}\left(\sum_{a=1}^{N-1}\alpha_{a}\left(E_{Na}-E_{NN}\right)\right)\exp{(E)}=\sum_{a=1}^{N-1}\left(E_{Na}-E_{aa}-\sum_{b=1\,b\neq a}^{N-1}E_{ba}\right)}
\fraT{and}

@@

	\fraT{\exp{(-E)}\left(\alpha_{N}\sum_{b=1}^{N-1}\alpha_{a}\left(E_{bN}-E_{bb}\right)\right)\exp{(E)}=-\alpha_{N}\sum_{b=1}^{N-1}E_{bb}}

@@

By using \fraT{\eqref{uno},\eqref{due} and \eqref{tre})} and the fact that 
\fraT{$\sum_{A,B=1}^{N}E_{AB}$} is central for the algebra we have

@@

\fraT{\sum_{a=1}^{N-1}\alpha_{a}\left(E_{Na}^{x}-E_{aa}^{x}\right)-\sum_{a,b=1\, a\neq b}^{N-1}\alpha_{a}^{x}E_{bb}^{x}-\alpha_{N}\sum_{a=1}^{N-1}E_{aa}^{x}\nonumber}

@@

\fraT{	|\alpha^{x}|\sum_{a=1}^{N-1}\left(\rho_{a}^{x}E_{Na}^{x}-E_{aa}^{x}\right)}\dt

@@

we obtain \fraT{

@@

{With a slight abuse of notation and in the spirit of Remark \ref{reamrk-extraSite}, we here denote the occupation vector at extra-site $u(x)$ by $\langle \xi_{1}^{u(x)},\ldots,\xi_{N-1}^{u(x)}|$\fraT{.}

@@

\fraT{Recalling} the action of the bosonic creation operator acting at site $u(x)$ and on the species $a\in \{1,\ldots,N-1\}$ we have that 

@@

Using the above equation \fraT{and considering the addends multiplied by the Lie generator $E_{Na}^{x}$ (for all $a\in \{1,\ldots,N-1\}$) on the right hand side term of \eqref{siteHadamardII}} we have that 

@@

In the long-time limit the dual process voids the chain, i.e. all particles of types $\{1,\ldots,N-1\}$ are eventually absorbed at the extra-sites $0$ and $L+1$ and replaced by types $N$. We also notice that, \fraT{up to the the bosonic} creation operators $(\mathbf{a}^{+})_{a}$, the dual boundary Hamiltonians \eqref{boundary-H-dual-chain} are triangular. This simplification will be crucial in Section~\ref{sectionIntegrabiliy}.\\

@@

\section{Correlations and steady state  for the integrable chain}\label{sectionIntegrabiliy}
In this section we combine the matrix product ansatz (MPA) \cite{vanicat2017exact} and  the $gl(N)$ invariance of the process in the bulk, cf.~\eqref{symmetryH}, to exactly compute the correlations in the non-equilibrium steady state
 for the integrable version of the multispecies stirring process \fraT{defined on a line segment}, see Section~\ref{integrableChain-duality}. 
This allows further to determine the absorption probabilities and the probability mass-function of the non-equilibrium steady state in closed form.
We start by recalling the formulation of the MPA  and then we explain our strategy.
